

---
title: "2008 vs. 2012 Presidential Election Analysis"
output: html_document
---
**Contributors:** Jacob Martella, Paul Freeman, Paul Vonderhaar

Here is the summary of the data:
```{r, echo=FALSE}
require("jsonlite")
require("RCurl")
# Change the USER and PASS below to be your UTEid
df1 <- data.frame(fromJSON(getURL(URLencode(gsub("\n", " ", '129.152.144.84:5001/rest/native/?query= "SELECT * FROM COPY_OF_2008PRESREFORMATTED;"')),httpheader=c(DB='jdbc:oracle:thin:@129.152.144.84:1521/PDBF15DV.usuniversi01134.oraclecloud.internal', USER='cs329e_jam8789', PASS='orcl_jam8789', MODE='native_mode', MODEL='model', returnDimensions = 'False', returnFor = 'JSON'), verbose = TRUE)));

df2 <- data.frame(fromJSON(getURL(URLencode(gsub("\n", " ", '129.152.144.84:5001/rest/native/?query="SELECT * FROM US_ELECTION_RESULTS;"')),httpheader=c(DB='jdbc:oracle:thin:@129.152.144.84:1521/PDBF15DV.usuniversi01134.oraclecloud.internal', USER='cs329e_jam8789', PASS='orcl_jam8789', MODE='native_mode', MODEL='model', returnDimensions = 'False', returnFor = 'JSON'), verbose = TRUE)));
summary(df1)
summary(df2)
```

Here is the head of our data:
```{r, echo=FALSE}
require("jsonlite")
require("RCurl")
# Change the USER and PASS below to be your UTEid
df1 <- data.frame(fromJSON(getURL(URLencode(gsub("\n", " ", '129.152.144.84:5001/rest/native/?query= "SELECT * FROM COPY_OF_2008PRESREFORMATTED;"')),httpheader=c(DB='jdbc:oracle:thin:@129.152.144.84:1521/PDBF15DV.usuniversi01134.oraclecloud.internal', USER='cs329e_jam8789', PASS='orcl_jam8789', MODE='native_mode', MODEL='model', returnDimensions = 'False', returnFor = 'JSON'), verbose = TRUE)));

df2 <- data.frame(fromJSON(getURL(URLencode(gsub("\n", " ", '129.152.144.84:5001/rest/native/?query="SELECT * FROM US_ELECTION_RESULTS;"')),httpheader=c(DB='jdbc:oracle:thin:@129.152.144.84:1521/PDBF15DV.usuniversi01134.oraclecloud.internal', USER='cs329e_jam8789', PASS='orcl_jam8789', MODE='native_mode', MODEL='model', returnDimensions = 'False', returnFor = 'JSON'), verbose = TRUE)));
head(df1)
head(df2)
```

**Finding Percent Change From 2008 to 2012**
One of the the things I was most interested in this data was how much the turnout changed between 2008 and 2012. My hypothesis was that turnout dropped in 2012 because there wasn't quite the same enthusiasm there was in 2008. But I was wrong. The data shows that turnout increased dramatically. Now, I think that there's an error between the two and that the 2012 data is probably more detailed and better done, accounting for a good chunk of the increase. But it's still an interesting part of the data to explore.


```{r}
require("jsonlite")
require("RCurl")
source('../02 Data Wrangling/PercentChangeWrangling.R', echo = TRUE)
```


```{r}
require("jsonlite")
require("RCurl")
source('../03 Data Visualization/PercentChangeVisualization.R', echo = TRUE)
```


**Reworking the data**
We suspected that one of the causes for the unreal information found in the previous graph is the original state of the graph. We found it on a government site on an excel sheet, and tried to just save that as a csv and use that data straight. This proved difficult, so we went into the data to try and modify it to get rid of some of the unnecessary information.
Here are the heads of our reformed data tables
```{r}
require("jsonlite")
require("RCurl")
df2008 <- data.frame(fromJSON(getURL(URLencode('129.152.144.84:5001/rest/native/?query="select * from ELECTIONOF2008 where CANDIDATE is not null"'),httpheader=c(DB='jdbc:oracle:thin:@129.152.144.84:1521/PDBF15DV.usuniversi01134.oraclecloud.internal', USER='cs329e_pmv347', PASS='orcl_pmv347', MODE='native_mode', MODEL='model', returnDimensions = 'False', returnFor = 'JSON'), verbose = TRUE), ))
head(df2008)
df2012 <- data.frame(fromJSON(getURL(URLencode('129.152.144.84:5001/rest/native/?query="select * from ELECTIONOF2012 where STATE is not null"'),httpheader=c(DB='jdbc:oracle:thin:@129.152.144.84:1521/PDBF15DV.usuniversi01134.oraclecloud.internal', USER='cs329e_pmv347', PASS='orcl_pmv347', MODE='native_mode', MODEL='model', returnDimensions = 'False', returnFor = 'JSON'), verbose = TRUE), ))
head(df2012)

head(df2012)
head(df2008)
```


**Wrangling the New Data**
Once we've gotten this data, we need to do a few things
First of all we need to turn all the vote counts of df 2012 into actual numbers,
we'd also like to get a total number of votes for each state, which we will need for later analysis
```{r}
require("jsonlite")
require("RCurl")
df2008 <- data.frame(fromJSON(getURL(URLencode('129.152.144.84:5001/rest/native/?query="select * from ELECTIONOF2008 where CANDIDATE is not null"'),httpheader=c(DB='jdbc:oracle:thin:@129.152.144.84:1521/PDBF15DV.usuniversi01134.oraclecloud.internal', USER='cs329e_pmv347', PASS='orcl_pmv347', MODE='native_mode', MODEL='model', returnDimensions = 'False', returnFor = 'JSON'), verbose = TRUE), ))
head(df2008)
df2012 <- data.frame(fromJSON(getURL(URLencode('129.152.144.84:5001/rest/native/?query="select * from ELECTIONOF2012 where STATE is not null"'),httpheader=c(DB='jdbc:oracle:thin:@129.152.144.84:1521/PDBF15DV.usuniversi01134.oraclecloud.internal', USER='cs329e_pmv347', PASS='orcl_pmv347', MODE='native_mode', MODEL='model', returnDimensions = 'False', returnFor = 'JSON'), verbose = TRUE), ))
head(df2012)

df2012$VOTESD<-as.numeric(gsub(",","",as.character(df2012$DEMOCRATVOTES)))
df2012$TOTPOP<-as.numeric(gsub(",","",as.character(df2012$TOTALPOPULATION)))
df2012$VOTESI<-as.numeric(gsub(",","",as.character(df2012$OTHERVOTES)))
df2012$VOTESD<-as.numeric(gsub(",","",as.character(df2012$DEMOCRATVOTES)))
df2012$VOTESR<-as.numeric(gsub(",","",as.character(df2012$REPUBLICANVOTES)))
df2012<-dplyr::mutate(df2012,PercentVoters=as.numeric((VOTESI+VOTESR+VOTESD)*100/TOTPOP))
```

After that, we start working with df2008
It has some strange quirks, so first we take the commas out of the vote number
After that we define a new dataframe, finalDF so we can put the data of df2008 into more useful collums
We also decided to add the State population to this table, which we'll use for calculation later
(Note, we removed the District of Columbia because it was no in the other data set, and We removed New York because three different values were given in the original
Data file)
```{r}
require("jsonlite")
require("RCurl")
df2008 <- data.frame(fromJSON(getURL(URLencode('129.152.144.84:5001/rest/native/?query="select * from ELECTIONOF2008 where CANDIDATE is not null"'),httpheader=c(DB='jdbc:oracle:thin:@129.152.144.84:1521/PDBF15DV.usuniversi01134.oraclecloud.internal', USER='cs329e_pmv347', PASS='orcl_pmv347', MODE='native_mode', MODEL='model', returnDimensions = 'False', returnFor = 'JSON'), verbose = TRUE), ))
head(df2008)
df2012 <- data.frame(fromJSON(getURL(URLencode('129.152.144.84:5001/rest/native/?query="select * from ELECTIONOF2012 where STATE is not null"'),httpheader=c(DB='jdbc:oracle:thin:@129.152.144.84:1521/PDBF15DV.usuniversi01134.oraclecloud.internal', USER='cs329e_pmv347', PASS='orcl_pmv347', MODE='native_mode', MODEL='model', returnDimensions = 'False', returnFor = 'JSON'), verbose = TRUE), ))
head(df2012)

df2012$VOTESD<-as.numeric(gsub(",","",as.character(df2012$DEMOCRATVOTES)))
df2012$TOTPOP<-as.numeric(gsub(",","",as.character(df2012$TOTALPOPULATION)))
df2012$VOTESI<-as.numeric(gsub(",","",as.character(df2012$OTHERVOTES)))
df2012$VOTESD<-as.numeric(gsub(",","",as.character(df2012$DEMOCRATVOTES)))
df2012$VOTESR<-as.numeric(gsub(",","",as.character(df2012$REPUBLICANVOTES)))
df2012<-dplyr::mutate(df2012,PercentVoters=as.numeric((VOTESI+VOTESR+VOTESD)*100/TOTPOP))

df2008$NUMVOTES<-as.numeric(gsub(",","",as.character(df2008$VOTES)))
finalDF<-data.frame(dplyr::filter(df2008,CANDIDATE=="Obama",STATE!="New York",STATE!="District of Columbia"))
finalDF$NumVotesR<-dplyr::filter(df2008,CANDIDATE=="McCain",STATE!="New York",STATE!="District of Columbia")$NUMVOTES
finalDF$TOTPOP<-as.numeric(filter(df2012,STATE!="New York")$TOTPOP)
finalDF<-dplyr::mutate(finalDF,PercentVoters2008=as.numeric((NumVotesR+NUMVOTES)*100/TOTPOP))

```

The headers for the datatables we are using now look like
```{r}
require("jsonlite")
require("RCurl")
df2008 <- data.frame(fromJSON(getURL(URLencode('129.152.144.84:5001/rest/native/?query="select * from ELECTIONOF2008 where CANDIDATE is not null"'),httpheader=c(DB='jdbc:oracle:thin:@129.152.144.84:1521/PDBF15DV.usuniversi01134.oraclecloud.internal', USER='cs329e_pmv347', PASS='orcl_pmv347', MODE='native_mode', MODEL='model', returnDimensions = 'False', returnFor = 'JSON'), verbose = TRUE), ))
head(df2008)
df2012 <- data.frame(fromJSON(getURL(URLencode('129.152.144.84:5001/rest/native/?query="select * from ELECTIONOF2012 where STATE is not null"'),httpheader=c(DB='jdbc:oracle:thin:@129.152.144.84:1521/PDBF15DV.usuniversi01134.oraclecloud.internal', USER='cs329e_pmv347', PASS='orcl_pmv347', MODE='native_mode', MODEL='model', returnDimensions = 'False', returnFor = 'JSON'), verbose = TRUE), ))
head(df2012)

df2012$VOTESD<-as.numeric(gsub(",","",as.character(df2012$DEMOCRATVOTES)))
df2012$TOTPOP<-as.numeric(gsub(",","",as.character(df2012$TOTALPOPULATION)))
df2012$VOTESI<-as.numeric(gsub(",","",as.character(df2012$OTHERVOTES)))
df2012$VOTESD<-as.numeric(gsub(",","",as.character(df2012$DEMOCRATVOTES)))
df2012$VOTESR<-as.numeric(gsub(",","",as.character(df2012$REPUBLICANVOTES)))
df2012<-dplyr::mutate(df2012,PercentVoters=as.numeric((VOTESI+VOTESR+VOTESD)*100/TOTPOP))

df2008$NUMVOTES<-as.numeric(gsub(",","",as.character(df2008$VOTES)))
finalDF<-data.frame(dplyr::filter(df2008,CANDIDATE=="Obama",STATE!="New York",STATE!="District of Columbia"))
finalDF$NumVotesR<-dplyr::filter(df2008,CANDIDATE=="McCain",STATE!="New York",STATE!="District of Columbia")$NUMVOTES
finalDF$TOTPOP<-as.numeric(filter(df2012,STATE!="New York")$TOTPOP)
finalDF<-dplyr::mutate(finalDF,PercentVoters2008=as.numeric((NumVotesR+NUMVOTES)*100/TOTPOP))

head(df2012)
head(finalDF)
```
Now that we've got these two sets of data organzied by state, we can join them
```{r}
require("jsonlite")
require("RCurl")
df2008 <- data.frame(fromJSON(getURL(URLencode('129.152.144.84:5001/rest/native/?query="select * from ELECTIONOF2008 where CANDIDATE is not null"'),httpheader=c(DB='jdbc:oracle:thin:@129.152.144.84:1521/PDBF15DV.usuniversi01134.oraclecloud.internal', USER='cs329e_pmv347', PASS='orcl_pmv347', MODE='native_mode', MODEL='model', returnDimensions = 'False', returnFor = 'JSON'), verbose = TRUE), ))
head(df2008)
df2012 <- data.frame(fromJSON(getURL(URLencode('129.152.144.84:5001/rest/native/?query="select * from ELECTIONOF2012 where STATE is not null"'),httpheader=c(DB='jdbc:oracle:thin:@129.152.144.84:1521/PDBF15DV.usuniversi01134.oraclecloud.internal', USER='cs329e_pmv347', PASS='orcl_pmv347', MODE='native_mode', MODEL='model', returnDimensions = 'False', returnFor = 'JSON'), verbose = TRUE), ))
head(df2012)

df2012$VOTESD<-as.numeric(gsub(",","",as.character(df2012$DEMOCRATVOTES)))
df2012$TOTPOP<-as.numeric(gsub(",","",as.character(df2012$TOTALPOPULATION)))
df2012$VOTESI<-as.numeric(gsub(",","",as.character(df2012$OTHERVOTES)))
df2012$VOTESD<-as.numeric(gsub(",","",as.character(df2012$DEMOCRATVOTES)))
df2012$VOTESR<-as.numeric(gsub(",","",as.character(df2012$REPUBLICANVOTES)))
df2012<-dplyr::mutate(df2012,PercentVoters=as.numeric((VOTESI+VOTESR+VOTESD)*100/TOTPOP))

df2008$NUMVOTES<-as.numeric(gsub(",","",as.character(df2008$VOTES)))
finalDF<-data.frame(dplyr::filter(df2008,CANDIDATE=="Obama",STATE!="New York",STATE!="District of Columbia"))
finalDF$NumVotesR<-dplyr::filter(df2008,CANDIDATE=="McCain",STATE!="New York",STATE!="District of Columbia")$NUMVOTES
finalDF$TOTPOP<-as.numeric(filter(df2012,STATE!="New York")$TOTPOP)
finalDF<-dplyr::mutate(finalDF,PercentVoters2008=as.numeric((NumVotesR+NUMVOTES)*100/TOTPOP))

dfVisualizer<-dplyr::inner_join(finalDF,df2012, by="STATE")
dfVisualizer<-dplyr::mutate(dfVisualizer,PercentVoters2008=as.numeric((NumVotesR+NUMVOTES)*100/TOTPOP.y))
```
We also took the liberty of calculating the perscent of voters in 2008 here

Now we want to start visualization using dfVisualizer
**Visualization using the new data**
We wanted to get 2 graphs at first with this data. 
First we wanted to compare the election results between the two years, then we wanted to look at the voter particpation per state
Those Graphs are below.
```{r}
require("jsonlite")
require("RCurl")
source('../03 Data Visualization/VonderHaarsDataVisualization.R', echo = TRUE)
```

These graphs have several interesting, nonobvious components. First of all, it seems that on the whole, the Democrats either experienced a slight increase
A sharp decline, or held steady in the polls. There were no states where they had very sharp increases. This is likely because Obama, as an incumbant,
had lost a lot of his initial luster as a candidate. On the other hand, the republicans made some notisable increases in a few states, implying
a return of those states to their natural political color. However, we can't confirm this without seeing at least a few more election results, and that is beyond the scope
of this process.
Overall, it seems like voter particpation is slightly down on this graph.

The second graph aslo has some interesting stories to tell. In very populace states such as California or Texas, Voter particpation is at its lowest.
If we compare the two graphs, its appears that neither of those states had "close" elections. We can thus say that Low voter particpation and Blowout elections 
are correlated, althought we can't say which causes which. I would guess that blow out elections cause lowe particpation, as people feel that their vote does not count,
but I do not have hard evidence to support this. In the same way, Battleground states like Florida and Ohio have very high voter particatpion, further strengthening
the idea of correlation. Again though, more years would allow us to draw better conclusions. I'm very interested what the particpation of say Nixon's re-election,
where he won 48 states, was compared to the average election.