---
title: "2008 vs. 2012 Presidential Election Analysis"
output: html_document
---
**Contributors:** Jacob Martella, Paul Freeman, Paul Vonderhaar

Here is the summary of the data:
```{r, echo=FALSE}
require("jsonlite")
require("RCurl")
# Change the USER and PASS below to be your UTEid
df1 <- data.frame(fromJSON(getURL(URLencode(gsub("\n", " ", '129.152.144.84:5001/rest/native/?query= "SELECT * FROM COPY_OF_2008PRESREFORMATTED;"')),httpheader=c(DB='jdbc:oracle:thin:@129.152.144.84:1521/PDBF15DV.usuniversi01134.oraclecloud.internal', USER='cs329e_jam8789', PASS='orcl_jam8789', MODE='native_mode', MODEL='model', returnDimensions = 'False', returnFor = 'JSON'), verbose = TRUE)));

df2 <- data.frame(fromJSON(getURL(URLencode(gsub("\n", " ", '129.152.144.84:5001/rest/native/?query="SELECT * FROM US_ELECTION_RESULTS;"')),httpheader=c(DB='jdbc:oracle:thin:@129.152.144.84:1521/PDBF15DV.usuniversi01134.oraclecloud.internal', USER='cs329e_jam8789', PASS='orcl_jam8789', MODE='native_mode', MODEL='model', returnDimensions = 'False', returnFor = 'JSON'), verbose = TRUE)));
summary(df1)
summary(df2)
```

Here is the head of our data:
```{r, echo=FALSE}
require("jsonlite")
require("RCurl")
# Change the USER and PASS below to be your UTEid
df1 <- data.frame(fromJSON(getURL(URLencode(gsub("\n", " ", '129.152.144.84:5001/rest/native/?query= "SELECT * FROM COPY_OF_2008PRESREFORMATTED;"')),httpheader=c(DB='jdbc:oracle:thin:@129.152.144.84:1521/PDBF15DV.usuniversi01134.oraclecloud.internal', USER='cs329e_jam8789', PASS='orcl_jam8789', MODE='native_mode', MODEL='model', returnDimensions = 'False', returnFor = 'JSON'), verbose = TRUE)));

df2 <- data.frame(fromJSON(getURL(URLencode(gsub("\n", " ", '129.152.144.84:5001/rest/native/?query="SELECT * FROM US_ELECTION_RESULTS;"')),httpheader=c(DB='jdbc:oracle:thin:@129.152.144.84:1521/PDBF15DV.usuniversi01134.oraclecloud.internal', USER='cs329e_jam8789', PASS='orcl_jam8789', MODE='native_mode', MODEL='model', returnDimensions = 'False', returnFor = 'JSON'), verbose = TRUE)));
head(df1)
head(df2)
```

**Finding Percent Change From 2008 to 2012**
One of the the things I was most interested in this data was how much the turnout changed between 2008 and 2012. My hypothesis was that turnout dropped in 2012 because there wasn't quite the same enthusiasm there was in 2008. But I was wrong. The data shows that turnout increased dramatically. Now, I think that there's an error between the two and that the 2012 data is probably more detailed and better done, accounting for a good chunk of the increase. But it's still an interesting part of the data to explore.


```{r}
require("jsonlite")
require("RCurl")
source('../02 Data Wrangling/PercentChangeWrangling.R', echo = TRUE)
```


```{r}
require("jsonlite")
require("RCurl")
source('../03 Data Visualization/PercentChangeVisualization.R', echo = TRUE)
```
